---
title: "PA1_template"
author: "Aleksei Dudin"
date: "August 9, 2018"
output:
  md_document: default
  'md_document: default': default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hi, lets start the peer-graded assignment!

### Load libraries

```{r, results = "hide"}
library(ggplot2)
library(knitr)
```

### Read the data

```{r}
path <- 'C:/Users/Aleksei Dudin/Documents/data1/activity.csv'
data <- read.csv(path)
```

### Check the data

```{r}
head(data)
```

```{r}
str(data)
```

```{r}
summary(data)
```

### What is mean total number of steps taken per day?

```{r, fig.height = 4, fig.width = 8}
steps.per.day <- tapply(data$steps, data$date, FUN = sum, na.rm = T)

qplot(steps.per.day, binwidth = 1000,
      color = I('red'),
      fill = I('white'),
      xlab='Total number of steps taken per day')
```

```{r}
mean(steps.per.day, na.rm = T)
```

```{r}
median(steps.per.day, na.rm = T)
```

```{r}
sum (steps.per.day, na.rm = T)
```

### What is the average daily activity pattern?

```{r, fig.height = 4, fig.width = 8}
averages <- aggregate(x = list(steps = data$steps),
                      by = list(interval = data$interval),
                      FUN = mean, na.rm = T)

ggplot(data = averages, aes(x = interval, y = steps)) +
    geom_line(col = 'red') +
    xlab('5-minute intervals') +
    ylab('Average number of steps taken')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averages[which.max(averages$steps),]
```

### Imputing missing values

Calculate and report the total number of missing values in the dataset.

```{r}
sum(is.na(data))
```

Use mean to fill in the `NA` values.

```{r}
fill.cells <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) filled <- c(steps) 
    else filled <- (averages[averages$interval == interval, 'steps'])
        return(filled)
}
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
filled.data <- data
filled.data$steps <- mapply(fill.cells, filled.data$steps,
                            filled.data$interval)
```

Make a histogram of the total number of steps taken each day.

```{r, fig.height = 4, fig.width = 8}
filled.steps.per.day <- tapply(filled.data$steps, filled.data$date, FUN = sum)
qplot(filled.steps.per.day, binwidth = 1000, colour = I('red'),
      fill = I('white'), xlab = 'Total number of steps taken per day')
```

 Calculate and report the mean and median total number of steps taken per day.
 
```{r}
mean(filled.steps.per.day)
median(filled.steps.per.day)
sum(filled.steps.per.day)
```

Remember the values without filled in `NA`.

```{r}
mean(steps.per.day, na.rm = T)
median(steps.per.day, na.rm = T)
sum (steps.per.day, na.rm = T)
```

Mean, median and total number steps values are higher after imputing missing data.

### Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday',
                   'Friday')) 
        return('weekday') else if (day %in% c('Saturday', 'Sunday')) 
            return('weekend') else stop('no date')
}
filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, fig.height = 4, fig.width = 8}
averages <- aggregate(steps ~ interval + day, data = filled.data, mean)
ggplot(averages, aes(interval, steps, color = day))+
    geom_line(show.legend=F) + facet_grid(day~.) + 
    xlab('5-minute intervals') + ylab('Number of steps')
```

## Thank you for reading!